### 2010-02-23 sarahDraft.tex
More writing...

\documentclass{article}
\usepackage{amsmath}
\begin{document}

@@

	Although we rejected the idea of using \(r\) as a way to analyze the data, we
realized that \(p\) and \(q\) could still be used to test the viability of the 
independence assumption.  If there is independece between sites, the difference 
of the proportions \(p\) and \(1-q\) should be approximately \(0\).  Looking at 
boxplots of \(p - (1-q)\), we see that data from flights that record alternating 
sites may be more independent than data from flights that record all sites, which
suggests a possible benefit to using the alternating sampling method.  The
boxplots for the Whitewater and Zumbro Rivers looked similar to the Mississippi:
*INSERT BOXPLOT FROM STATUS REPORT HERE
We found no obvious trend indicating one site size to be more independent than 
the others.

	While this is an interesting observation, we would like to be able to 
quantify it better by finding the distribution of our statistic.

		We wanted to find out if the assumption of independence across sites held
true for our data from the DNR.  We also wanted to see what sampling techniques 
led to the greatest independence between plots.  The different techniques that we
compared were using different plot sizes, and using data that includes all plots
or data that uses only every other plot. \\

		If the observations at different sites are independent, then the 
probability that there is a \(1\) at site \(i\)  is independent of site \(i-1\).  
We let \(p\) be the number of \(0\)'s followed by a \(1\) as a fraction of all 
successive pairs that start with a \(0\).  We let \(1-q\) be the number of \(1\)'s
followed by a \(1\) as a fraction of all successive pairs that start with a \(1\).
The difference of these two proportions \((=p-(1-q)\) is a measure of independence,
where \(p-(1-q)\) being approximately \(0\) under the null hypothesis of 
independence.  
	\begin{enumerate}
    	\item Using Markov Chains to find r. \\
          * r = potential long-term probability of detection, based upon the 
			short-term conditional probabilities p and q acquired from our data 
			(infinite limit as n approaches infinity of Pn) \\
          * p = probability that a 0 follows a 0, q = probability that a 1 follows
			 a 1  \\
          * If 0->0 and 1->0 (the values of a column, basically) equaled each 
			other, plots would have been independent?  \\
          * Ultimately not useful to us because it does not take into account 
			whether or not a site is occupied, so it doesn't apply to the 
			hierarchical model that we are using. \\
   		\item Using s, where s = (\# of 0-1's) - (\# of 1-1's) for a given flight.\\
          * Decided to use the z-score of s as a measure of independence.  \\
          * Therefore, needed the expected value and the standard deviation of s. \\
          * So, we needed the expected value of each term and the variance, which 
			required finding the variance of each term and the covariance between 
			the two terms. \\
          * Expected value is straightforward and we have all of the equations. \\
          * Variance is fun and we also have all of the relevant equations. \\
          * Calculated z-score for every flight and box-plotted them for each year
			and plot size and pitted data with every other plot against data with
			all plots. \\
          * Nothing important with respect to plot size. Alternating plots show 
			more independence than all plots. Look at report for more. \\
          * CAR model started working once we took s values of simulated data into
			account by increasing dependence.  \\
	\end{enumerate}

>>>>

\documentclass{article}
\usepackage{amsmath}
\usepackage{fullpage}
\begin{document}

@@

	Although we rejected the idea of using \(r\) as a way to analyze the data, we
realized that \(p\) and \(q\) could still be used to test the viability of the 
independence assumption.  If there is independece between sites, the difference 
of the proportions \(p\) and \(1-q\) should be approximately \(0\).  This would 
mean that the probability that a \(0\) follows a \(0\) is about the same as the 
probability that a \(1\) follows a \(0\), meaning that it does not depend on what
the observation is at the previous site.  Looking at boxplots of \(p - (1-q)\), 
we see that data from flights that record alternating sites may be more 
independent than data from flights that record all sites, which suggests a 
possible benefit to using the alternating sampling method.  The boxplots for the 
Whitewater and Zumbro Rivers looked similar to the Mississippi:  \\
*INSERT BOXPLOT FROM STATUS REPORT HERE  \\
We found no obvious trend indicating one site size to be more independent than 
the others.

	While this is an interesting observation, we would like to be able to 
quantify it better by finding the distribution of our statistic.  This would 
allow us to make statements about how independent the data is, and if the 
difference in independence between different types of data is significant.   
This statistic seems hard to get, but we considered a modification.  Let \(S\) be
the number of successive \(0\),\(1\)'s minus the number of successive \(1\),\(1\)'s
per flight.  Under the null hypothesis that each site is an independent Bernoulli
trial with probability \(p\) we were able to compute the mean and standard 
deviation of \(S\).

	The mean ...

	The standard deviation...

	Simulations of \(S\) show that the distribution of \(S\) is roughly normal.  
So using our calculations for the mean and the standard deviation, we computed
a Z-score for each flight, where we use a plug-in estimate for \(p\) to be the 
proportion of \(1\)'s per flight.  We said that Z-scores with an absolute value 
greater than \(2\) suggest that the assumption of independence between sites is 
violated for that flight.  We found that, in general, alternating sites are more
"independent" than all sites as measured by this statistic.  However, alternating
sites do not guarentee independence.  For instance, six of eleven flights along 
the Zumbro River during 2003 still have significant Z-scores with alternating
sites.  There were no noticable trends for this statistic with respect to 
observer, days since snow, or snow event.

*FIGURES OR TABLES??? WHICH ONES?

	We did find interesting trends with regard to sampling method.  We think that 
the lack of independence results from strings of ones where an otter left a track
across multiple sites.  When alternating sites are used, the strings are cut in 
half and the occurrence of \(1\),\(1\)'s is lower while the occurence of \(0\),
\(1\)'s stays about the same.  It is noteworthy that virtually all the violations
of independence are in the direction of more \(1\),\(1\)'s than would be expected.
This reinforces our thoughts about successive \(1\)'s representing the same otter.
We will discuss later whether this violation of independence actually matters in 
our model, and how this might affect the estimation of actual occupancy rates.

   		Using s, where s = (\# of 0-1's) - (\# of 1-1's) for a given flight.\\
          * Decided to use the z-score of s as a measure of independence.  \\
          * Therefore, needed the expected value and the standard deviation of s. \\
          * So, we needed the expected value of each term and the variance, which 
			required finding the variance of each term and the covariance between 
			the two terms. \\
          * Expected value is straightforward and we have all of the equations. \\
          * Variance is fun and we also have all of the relevant equations. \\
          * Calculated z-score for every flight and box-plotted them for each year
			and plot size and pitted data with every other plot against data with
			all plots. \\
          * Nothing important with respect to plot size. Alternating plots show 
			more independence than all plots. Look at report for more. \\
